### 2018-03-07 latex/sc18/abstract.tex
splitting sections

\begin{abstract}
The abstract goes here.
\end{abstract}