### 2018-03-07 latex/sc18/introduction.tex

\section{Introduction}

\grey{Contributions:
\begin{itemize}
	\item fastest scalable NNCP (and unsonstrained CP) software for dense tensors
	\item computation efficient using dimension trees, communication efficient using optimal parallel algorithm
	\item available open-source software, with flexibility to incorporate multiple NLS algorithms
\end{itemize}
}