### 2018-03-07 latex/sc18/prelims.tex
added lists after conversation

\section{Preliminaries} 
\label{sec:prelims}

\begin{itemize}
	\item tensor notation, Hadamard, Khatri-Rao, MTTKRP
	\item NNCP definition
	\item BCD/ALS
	\begin{itemize}
		\item MU, HALS, BPP, Nesterov, ADMM
	\end{itemize}
	\item MPI collectives
\end{itemize}