### 2018-03-07 latex/sc18/survey.tex

\section{Related Work} 
\label{sec:survey}

\grey{I haven't organized these very well...}
\begin{itemize}
	\item NLS algorithms
	\begin{itemize}
		\item MU, HALS, BPP, etc.
	\end{itemize}
	\item parallel sparse CP
	\begin{itemize}
		\item Karypis/Smith
		\item Kaya/Ucar
		\item Vagelis/Faloutsos?
		\item Li/Vuduc?
	\end{itemize}
	\item dense tensor computations
	\begin{itemize}
		\item Phan/Cichocki
		\item Austin/Ballard/Kolda IPDPS16
		\item Ballard/Knight/Rouse IPDPS18
	\end{itemize}
	\item parallel dense CP
	\begin{itemize}
		\item Liavs et al.
		\item Cichocki?
		\item distributed incremental descent?
	\end{itemize} 
\end{itemize}