### 2018-03-20 latex/sc18/abstract.tex
first draft of abstract

% !TEX root = paper.tex

\begin{abstract}

The CP tensor decomposition is a low-rank approximation of multidimensional (or tensor) data. 
We present a distributed-memory parallel algorithm and implementation of an alternating optimization method for computing a CP decomposition of dense tensor data that can enforce nonnegativity of the computed low-rank factors.
The principal task is to parallelize the matricized-tensor times Khatri-Rao product (MTTKRP) bottleneck subcomputation.
The algorithm is computation efficient, using dimension trees to avoid redundant computation across MTTKRPs within the alternating method.
Our approach is also communication efficient, using a data distribution and parallel algorithm across a multidimensional processor grid that can be tuned to minimize communication.
We benchmark our software on synthetic as well as hyperspectral image data, demonstrating that our algorithm scales well to 100s of nodes (up to 9600 cores) and is faster than the currently available parallel software.

\end{abstract}