### 2018-03-22 latex/sc18/plots.tex
starting to put together plots

% macros for plotting
\newcommand{\datafile}{}
\newcommand{\alg}{}
%\newcommand{\numiterations}{30}
%\newcommand{\minvalue}{1}

% toggles whether or not to plot naive results (if not, the rows of the data file also need to be commented out)
\newif\ifnaive
% toggles ksweep vs scaling plot
\newif\ifksweep
% toggles legend
\newif\iflegend
% toggles y axis label
\newif\ifylabel
% toggles wider plot area
\newif\ifwider

\newcommand{\setcolors}{
\pgfplotsset{cycle list={
	red, fill=red \\ 
	blue, fill=blue \\ 
	green, fill=green \\ 
	red, pattern=crosshatch, pattern color=red \\
	blue, pattern=crosshatch, pattern color=blue \\
	green, pattern=crosshatch, pattern color=green \\
}};
}

% set options for grouped bar plot
\newcommand{\plotoptions}{
	ybar stacked,
	reverse legend,
	bar width=6pt,
	width=9cm, height=3.85cm,
	ylabel={Time (seconds)}, 
	y label style={yshift=-.5cm},
	ymin=0,
	\ifnaive 
		\ifksweep
			symbolic x coords={10-0,10-1,10-2,10-3,,20-0,20-1,20-2,20-3,,30-0,30-1,30-2,30-3,,40-0,40-1,40-2,40-3,,50-0,50-1,50-2,50-3},
		\else
			symbolic x coords={16-0,16-1,16-2,16-3,,96-0,96-1,96-2,96-3,,384-0,384-1,384-2,384-3,,864-0,864-1,864-2,864-3,,1536-0,1536-1,1536-2,1536-3},
		\fi
		xticklabels={MU,HALS,ABPP,Naive,MU,HALS,ABPP,Naive,MU,HALS,ABPP,Naive,MU,HALS,ABPP,Naive,MU,HALS,ABPP,Naive},
	\else  
		\ifksweep
			symbolic x coords={10-0,10-1,10-2,,,20-0,20-1,20-2,,,30-0,30-1,30-2,,,40-0,40-1,40-2,,,50-0,50-1,50-2,},
		\else
			symbolic x coords={16-0,16-1,16-2,,,96-0,96-1,96-2,,,384-0,384-1,384-2,,,864-0,864-1,864-2,,,1536-0,1536-1,1536-2,}, 
		\fi
		xticklabels={MU,HALS,ABPP,MU,HALS,ABPP,MU,HALS,ABPP,MU,HALS,ABPP,MU,HALS,ABPP},
	\fi
	xtick=data,
	xticklabel style={xshift=.15cm,rotate=75,anchor=east},
	\ifksweep
		xlabel={Low Rank ($k$)}, 
	\else
		xlabel={Number of Processes ($p$)},
	\fi
	xlabel style={yshift=-0.75cm},
	%legend style={at={(0.4,-0.9)},anchor=north},
	legend style={at={(0.5,1.3)},anchor=north},
	legend columns=-1,
	legend style={draw=none, cells={align=left}, nodes={scale=0.8}}
}

% grouped bar plot command
\newcommand{\makeplot}{
\begin{axis}[\plotoptions]
	\setcolors
	\addplot table[x=algo, y expr=(\thisrow{mttkrp}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{nnls}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{gram}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{allgather}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{reducescatter}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{allreduce}/(\minvalue*\numiterations))] {\datafile};
	\iflegend
		\legend{MM, \NLS, Gram, All-Gather, Reduce-Scatter, All-Reduce};
	\fi
\end{axis}
}

% labels for bar groups (manually positioned)
\newcommand{\labels}{
%\node [align=center,text width=3cm] at (1.25cm, -.95cm)   {\ifksweep 10 \else 16 \fi};
%\node [align=center,text width=3cm] at (3cm, -0.95cm)   {\ifksweep 20 \else 96 \fi};
%\node [align=center,text width=3cm] at (4.6cm, -0.95cm)   {\ifksweep 30 \else 384 \fi};
%\node [align=center,text width=3cm] at (6.3cm, -0.95cm) {\ifksweep 40 \else 864 \fi};
%\node [align=center,text width=3cm] at (8.15cm, -0.95cm) {\ifksweep 50 \else 1536 \fi};
\node [align=center,text width=3cm] at (1cm, -1.15cm)   {\ifksweep 10 \else 16 \fi};
\node [align=center,text width=3cm] at (2.4cm, -1.15cm)   {\ifksweep 20 \else 96 \fi};
\node [align=center,text width=3cm] at (3.7cm, -1.15cm)   {\ifksweep 30 \else 384 \fi};
\node [align=center,text width=3cm] at (5.1cm, -1.15cm) {\ifksweep 40 \else 864 \fi};
\node [align=center,text width=3cm] at (6.4cm, -1.15cm) {\ifksweep 50 \else 1536 \fi};
}

% allows for filtering rows from dat file
\pgfplotsset{
    discard if not/.style 2 args={
        x filter/.append code={
            \edef\tempa{\thisrow{#1}}
            \edef\tempb{#2}
            \ifx\tempa\tempb
            \else
                \def\pgfmathresult{inf}
            \fi
        }
    }
}

% set options for strongscaling plot
\newcommand{\strongscalingplotoptions}{
	xlabel=Time (s),
	\ifylabel
		ylabel=Relative Error,
		ylabel style={yshift=-.3cm},
	\fi
	\ifwider
		width=1.8in, height=2in,
	\else
		width=1.35in, height=2in,
	\fi
	y tick label style={
        /pgf/number format/.cd,
            precision=4,
	},
	legend style={draw=none, cells={align=left}, nodes={scale=0.7}}
}

% plot time over p, filtering rows out appropriately
\newcommand{\strongscalingplot}{
\begin{axis}[\strongscalingplotoptions]
	%\addplot+ [discard if not={algo}{liavas},thick,mark options={solid}] table [x={nodes}, y={totaltime}] {\datafile};
	%\addplot+ [discard if not={algo}{nodimtree},thick,mark options={solid},mark=square*] table [x={nodes}, y={totaltime}] {\datafile};
	\addplot+ [discard if not={alg}{D},thick,mark options={solid},mark=triangle*] table [x={p}, y={total_d}] {\datafile};
	\iflegend
		\legend{Liavas, Classic, DimTree}
	\fi
\end{axis}
}