### 2018-03-29 latex/sc18/conclusion.tex

\section{Conclusion} \label{sec:conclusion}
\begin{itemize}
	\item We present new parallel software for distributed memory NNCP.
	\item We demonstrate consistent speed up against existing state of the art up to $2.2\times$.
	\item We extend our algorithms and software to a general number of modes.
	\item We present analysis of our parallel algorithms showing good strong and weak parallel scaling up to thousands of cores (hundreds of nodes)
\end{itemize}