### 2018-04-02 sc17/plots.tex

% macros for plotting
\newcommand{\datafile}{}
\newcommand{\numiterations}{30}
\newcommand{\minvalue}{1}
%\newcommand{\xticklabels}{\unp, \urp, \crp, \cpp}

\newif \ifylabel
\newif \iflegend
\newif \ifbgq

\newcommand{\setcolors}{
\pgfplotsset{cycle list={
	red, fill=red \\ 
	blue, fill=blue \\ 
	green, fill=green \\ 
	red, pattern=crosshatch, pattern color=red \\
	blue, pattern=crosshatch, pattern color=blue \\
	green, pattern=crosshatch, pattern color=green \\
}};
}
% set options for grouped bar plot
\newcommand{\plotoptions}{
	ybar stacked,
	reverse legend,
	bar width=8pt,
	width=2in, height=2in,
	\ifylabel
	     ylabel={Time (seconds)}, 
	     y label style={yshift=-.5cm},
	\fi
         ymin=0,
	%xticklabels/.expanded={\unp, \urp, \crp, \cnp}, 
	symbolic x coords={un, ur, cr, cp},
	xtick=data,
	xticklabel style={xshift=.15cm,rotate=45,anchor=east},
	%xlabel={Partitioning Schemes},
	xlabel style={yshift=-0.5cm},
	legend cell align=left,
	legend style={nodes={scale=0.7, transform shape}, draw=none,row sep=-0.1cm},
	%legend style={at={(1,.5)},anchor=west}
	legend pos=north east
}

\newcommand{\plotoptionsnew}{
	ybar stacked,
	reverse legend,
	bar width=8pt,
	width=2in, height=2in,
	\ifylabel
	     ylabel={Time (seconds)}, 
	     y label style={yshift=-.5cm},
	\fi
         ymin=0,
	symbolic x coords={cp-fl,cr-fl,cp-dl,cr-dl},
	xtick=data,
	xticklabel style={xshift=.15cm,rotate=45,anchor=east},	
	xlabel style={yshift=-0.5cm},
	legend cell align=left,
	legend style={nodes={scale=0.7, transform shape}, draw=none,row sep=-0.1cm},
	%legend style={at={(1,.5)},anchor=west}
	legend pos=north west
}

\newcommand{\speedupoptions}{
xlabel=Processors, ylabel=Speed-up,width=3.5in, height=2.0in, 
y label style={yshift=-.5cm}, 
legend pos=north west, legend style={draw=none,row sep=-0.1cm},	
xtick=data, 
\ifbgq
xticklabels={256, 512, 1024, 2048, 4096, 8192, 16384, 32768},
symbolic x coords={256, 512, 1024, 2048, 4096, 8192, 16384, 32768}
\else
xticklabels={16, 48, 96, 192, 384, 768, 1536, 3072},
symbolic x coords={16, 48, 96, 192, 384, 768, 1536, 3072}
\fi
}

% grouped bar plot command
\newcommand{\makeplot}{
\begin{axis}[\plotoptions, xticklabels/.expand once={\cpp, \crp, \unp, \urp}]
	\setcolors
	\addplot table[x=algo, y expr=(\thisrow{mm}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{nnls}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{gram}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=((\thisrow{allgather}+\thisrow{reducescatter})/(\minvalue*\numiterations))] {\datafile};
	%\addplot table[x=algo, y expr=(\thisrow{reducescatter}/(\minvalue*\numiterations))] {\datafile};
	\iflegend
	    %\legend{MM, \NLS, Gram, All-Gather, Reduce-Scatter};
	    \legend{MM, \NLS, Gram, Comm};
	\fi
\end{axis}
}

\newcommand{\makeplotnew}{
\begin{axis}[\plotoptionsnew, xticklabels/.expand once={\cpp,\crp,\cpp,\crp},
xlabel/.expand once={Flickr \quad\quad\quad Delicious} ]
	\setcolors
	\addplot table[x=algo, y expr=(\thisrow{mm}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{nnls}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{gram}/(\minvalue*\numiterations))] {\datafile};
	%\addplot table[x=algo, y expr=(\thisrow{err}/(\minvalue*\numiterations))] {\datafile};
	\addplot table[x=algo, y expr=(\thisrow{comm}/(\minvalue*\numiterations))] {\datafile};
	\iflegend
	    %\legend{MM, \NLS, Gram, Err, Comm};
	    \legend{MM, \NLS, Gram, Comm};
	\fi
\end{axis}
}

% labels for bar groups (manually positioned)
\newcommand{\labels}{
\node [align=center,text width=3cm] at (1.25cm, -.95cm){\cp};
\node [align=center,text width=3cm] at (3cm, -0.95cm) {\cr};
\node [align=center,text width=3cm] at (4.6cm, -0.95cm){\un};
\node [align=center,text width=3cm] at (6.3cm, -0.95cm){\ur};
}

\newcommand{\speedup}{
\begin{axis}[\speedupoptions]
\addplot [red,very thick, mark=triangle] table [x={p}, y expr=(\minvalue/\thisrow{cr})] {\datafile};
\addplot [green,very thick, mark=x] table [x={p}, y expr=(\minvalue/\thisrow{cp})] {\datafile};
\ifbgq
\else
\addplot [very thick, mark=diamond] table [x={p}, y expr=(\minvalue/\thisrow{ur})] {\datafile};
\addplot [blue,very thick, mark=*] table [x={p}, y expr=(\minvalue/\thisrow{un})] {\datafile};
\fi
\legend{\crp, \cpp, \urp, \unp}
\end{axis}
}

\newcommand{\numkblocksweep}{

\begin{axis}[xlabel=Number of $k$ blocks, 
	ylabel=Sparse-Dense MM (in Secs),
	%xticklabels={1,2,3,4,6},
	y label style={yshift=-.25cm},
	width=3in, height=2.4in]
\addplot[red,very thick] table [x={numkblocks}, y expr=(\thisrow{mm}/(\minvalue*\numiterations))] {\datafile}; \label{plot_one}
\end{axis}
}