### 2018-06-20 latex/toms/bio.tex

\vspace{-0.5in}
\begin{IEEEbiography}[{\includegraphics[width=1in,height =2.25in,clip,keepaspectratio]{fig/ramki}}]{Ramakrishnan Kannan}
is a Computational Data Scientist in 
Oak Ridge National Laboratory.
% focusing on large scale data mining and machine learning algorithms on HPC systems and modern architectures with applications from scientific domain and many different internet services. 
He received his Ph.D in Computer Science from College of Computing, 
Georgia Institute of Technology advised by Prof. Haesun Park. 
He worked on Data Analytics group at IBM TJ Watson Research Center and 
was an IBM Master Inventor. He has M.Sc (Engg) from Indian Institute of Science.  
\end{IEEEbiography}

\vspace{-0.5in}
\begin{IEEEbiography}[{\includegraphics[width=1in,height =1.25in,clip,keepaspectratio]{fig/ballard}}]{Grey Ballard}
Grey Ballard is an Assistant Professor in the Computer Science Department at Wake Forest University. After receiving his PhD in computer science from the University of California Berkeley in 2013, he was a Truman Fellow at Sandia National Laboratories in Livermore, CA. 
He received his BS in math and computer science at Wake Forest in 2006 and his MA in math at Wake Forest in 2008.
%His research interests include numerical linear algebra, high performance computing, and computational science, particularly in developing algorithmic ideas that translate to improved implementations and more efficient software. 
His work has been recognized with the SIAM Linear Algebra Prize and three conference best paper awards, at SPAA, IPDPS, and ICDM.
%; the C.V. Ramamoorthy Distinguished Research Award at UC Berkeley; and the ACM Doctoral Dissertation Award - Honorable Mention.
\end{IEEEbiography}
\vspace{-0.5in}

\begin{IEEEbiography}[{\includegraphics[width=1in,height =1.25in,clip,keepaspectratio]{fig/hpark}}]{Haesun Park}
is a IEEE and SIAM Fellow and Professor in the 
School of Computational Science and Engineering, 
Georgia Institute of Technology. 
%Her research areas include numerical algorithms, data analysis, visual analytics, text mining, and parallel computing.
She has played major leadership  roles as the Executive Director
of the Center for Data Analytics, Georgia Tech, general chair for the SIAM 
Conference on Data Mining, and editorial board member of SIAM
and IEEE journals. She received a  Ph.D. in Computer Science 
from Cornell University in 1987 and B.S. in Mathematics from Seoul National 
University with the University President's Medal.
\end{IEEEbiography}