### 2018-06-20 latex/toms/stackxchangetopics.tex

\begin {tabular}{|ccccc||>{\columncolor [gray]{.8}}c>{\columncolor [gray]{.8}}c>{\columncolor [gray]{.8}}c>{\columncolor [gray]{.8}}c>{\columncolor [gray]{.8}}c|}%
\toprule 
\multicolumn{5}{|c||}{Top Keywords from Topics 1-25} & \multicolumn{5}{|>{\columncolor [gray]{0.8}}c|}{Top Keywords from Topics 26-50} \\ 
word1&word2&word3&word4&word5&word1&word2&word3&word4&word5\\\midrule %
refer&undefin&const&key&compil&echo&type=text&php&form&result\\%
text&field&box&word&static&test&perform&fail&unit&result\\%
imag&src&descript&alt=ent&size&tabl&key&queri&databas&insert\\%
button&click&event&form&add&user&email&usernam&login&log\\%
creat&bean&add&databas&except&data&json&store&read&databas\\%
string&static&final&catch&url&page&load&content&url&link\\%
width&height&color&left&display&privat&static&final&import&float\\%
app&applic&servic&thread&work&row&column&date&cell&valu\\%
ipsum&lorem&dolor&sit&amet&line&import&command&print&recent\\%
node&list&root&err&element&var&map&marker&match&url\\%
0x00&0xff&byte&0x01&0xc0&server&connect&client&messag&request\\%
file&directori&read&open&upload&number&byte&size&print&input\\%
function&call&event&work&variabl&object&properti&json&instanc&list\\%
int&char&const&static&doubl&array&element&valu&key&index\\%
public&overrid&virtual&static&extend&main&thread&program&frame&cout\\%
return&param&result&def&boolean&type&field&properti&argument&resolv\\%
info&thread&start&map&servic&select&item&queri&join&list\\%
error&syntax&found&symbol&fail&sourc&target&except&java&fail\\%
set&properti&virtual&default&updat&instal&version&packag&err&default\\%
case&break&switch&default&cout&code&work&problem&chang&write\\%
method&call&except&static&todo&void&overrid&protect&catch&extend\\%
href&nofollow&src&link&work&true&requir&boolean&option&valid\\%
end&def&dim&begin&properti&find&project&import&warn&referenc\\%
debug&request&filter&match&found&view&control&item&overrid&posit\\%
fals&boolean&fix&bool&autoincr&null&default&key&int(11&primari\\\bottomrule %
\end {tabular}%