### 2019-01-16 latex/slides/SIAM-ALA18/fig/modeled.tex

%!TEX root = ../talk.tex

\begin{tikzpicture}[scale=.8]
\renewcommand{\datafile}{data/modeled.dat}
\begin{axis}[
	title=Modeled Strong-Scaling Comparison,
	xlabel=Processors,
	ylabel=Words Communicated,
	width=\columnwidth,
	xmode=log,
	log basis x={2},
	xmin=1,xmax=2^31,
	xtick={1,32,1024,32768,1048576,33554432,1073741824},
	ymode=log,
	legend pos=south west
	]
	\addplot [black,thick,dashed,mark options={solid},mark=x] table [x={P}, y={MM}] {\datafile};
	\addlegendentry{Matrix Multiplication}
	\addplot [red,thick,dashed,mark options={solid},mark=asterisk] table [x={P}, y={Alg3}] {\datafile};
	\addlegendentry{Stationary Tensor}
	\addplot [blue,thick,dashed,mark options={solid},mark=o] table [x={P}, y={Alg4}] {\datafile};
	\addlegendentry{General Tensor}
\end{axis}
\end{tikzpicture}