### 2019-01-16 latex/toms/ack.tex

\section{Acknowledgements}
\small

This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. OAC-1642385.
This manuscript has been co-authored by UT-Battelle, LLC under Contract No. DE-AC05-00OR22725 with the U.S. Department of Energy.  This project was partially funded by the Laboratory Director's Research and Development fund. This research used resources of the Oak Ridge Leadership Computing Facility at the Oak Ridge National Laboratory, which is supported by the Office of Science of the U.S. Department of Energy.

%This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231.

Also, partial funding for this work was provided by AFOSR Grant FA9550-13-1-0100, National Science Foundation (NSF) grants IIS-1348152, ACI-1338745, and ACI-1642385, Defense Advanced Research Projects Agency (DARPA) XDATA program grant FA8750-12-2-0309.
%We also thank NSF for the travel grant to present this work in the conference through the grant CCF-1552229. 
%The United States Government retains and the publisher, by accepting the article for publication, acknowledges that the United States Government retains a non-exclusive, paid-up, irrevocable, world-wide license to publish or reproduce the published form of this manuscript, or allow others to do so, for United States Government purposes. The Department of Energy will provide public access to these results of federally sponsored research in accordance with the DOE Public Access Plan (\url{http://energy.gov/downloads/doepublic-access-plan}). 
 Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the USDOE, NERSC, AFOSR, NSF or DARPA.

 \normalsize